---
title: "My post for Tidytuesday"
description: |
  In this blogpost, I will be consolidating all the weekly posts I submit for tidytuesday which is an R learning community challenge whereby R learners across the world are given datasets (either clean or not clean) and they come up with informative plots/figures, etc.
author:
  - name: Birasafab
    url: {https://github.com/birasafab/Tidytuesday2022}
date: 2022-02-22
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---



# First submission: TidyTuesday WEEK 08 (22 FEBRUARY 2022) --------------------------------------------------------------------

```{r setup, include=FALSE}

# Load the required packages and weekly Data

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```


# Freedom in the world

In this week's tidytuesday, I have visualized the worldwide countries' freedom status and the submitted plot looks like the following:

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Loading this week's dataset
freedom = readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-22/freedom.csv')


# Data wrangling and plotting

freedom %>% 
  mutate(Status = ifelse(Status=="F","Free",ifelse(Status=="NF","Not Free",ifelse(Status == "PF", "Partially Free",Status)))) %>% 
  group_by(year) %>% 
  count(Status) %>% 
  mutate(Status = fct_reorder(Status,-n)) %>% 
  ggplot(aes(year,n, fill = Status))+
  geom_area()+
  # geom_label(aes(label = Status))+
  labs(
    title = "Number of countries per Freedom status",
    y = "Number of countries",
    x = "Year of observation",
    caption = "Plot by Birasafab | Freedom House & United Nations | #Tidytuesday",
    fill = "Countries' Freedom status",
    subtitle = "Many of the countries worldwide were reported to be free followed by \n partially free countries group and those which are not free"
  )+
  theme(
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "black"),
    panel.background = element_rect(fill = "black"),
    axis.title = element_text(color = "white"),
    plot.title = element_text(color = "white"),
    plot.subtitle = element_text(color = "white"),
    legend.background = element_rect(fill = "black"),
    legend.title = element_text(color ="white"),
    legend.text = element_text(color= "white"),
    plot.caption = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.ticks = element_line(color = "white"),
    axis.line = element_line(color = "white")
  )
```

If you are interested in the above plot, feel free to view its [source code](https://github.com/birasafab/Tidytuesday2022/tree/main/2022_week08)


# My second submission for Week 08 tidytuesday --------------------------------------------------------------------

## Level of freedom per continent
```{r echo=FALSE, message=FALSE, warning=FALSE}
freedom %>% 
  janitor::clean_names() %>% 
 # filter(country %in% c("Rwanda","Uganda","Burundi","Tanzania","Kenya")) %>% 
  group_by(region_name, year) %>% 
  mutate(IFW = (cl+pr)/2,
         averaged_freedom = mean(IFW)) %>% 
  ungroup() %>% 
  group_by(region_name,year) %>% 
  ggplot(aes(year,averaged_freedom, color = region_name))+
  geom_line(size = 2)+
  scale_color_manual(values = c("#ff66ff","#33ff99","#000000","#3399ff","#3333ff"))+
  bbplot::bbc_style()+
  scale_y_reverse()+
  expand_limits(y= 1)+
  labs(
    title = "Continental level of freedom",
    subtitle = "Asia and Africa are the least free continents \n while Ocenia and Europe are the most free continents",
    caption = "plot by Birasafab | Freedom House & UN data by way of Arthur Cheib"
  )+
 theme(
    plot.caption = element_text(color = "black", face = "bold", size = 10),
    plot.title = element_blank(),
    plot.subtitle = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 9)
  )
```

In the above plot, Africa and Asia were observed to the continents with the least freedom while Europe and Oceania were the continents with highest freedom. 